---
title: "Analysis – Replication of Schoonvelde et al. (2019)"
author: "Colton Stevenson"
date: "April 3rd 2026"
output:
  html_document:
    theme: darkly
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: true
    toc_depth: 5
    highlight: zenburn
---

## Overview

This notebook reproduces the main analyses from:

> Schoonvelde, M., Brosius, A., & van Atteveldt, W. (2019). Liberals lecture,
> conservatives communicate: Analyzing complexity and ideology in 381,609
> political speeches. *PLOS ONE*, 14(2), e0208450.

We replicate the descriptive statistics (Tables A.1-A.8), the four main figures,
and the OLS regression tables (with and without speaker fixed effects).

**Input:** `data/cleaned/cole_final_cleaned_data.RData` (produced by `cleaning.Rmd`)

---

## Setup

### Libraries

```{r load-libraries, message = FALSE, warning = FALSE}

library(tidyverse)          # data wrangling and plotting
library(tidylog)            # verbose dplyr output
library(tidytext)           # tidy text tools
library(textdata)           # text datasets
library(quanteda)           # corpus tools
library(quanteda.textstats) # readability
library(wesanderson)        # colour palettes
library(knitr)              # table formatting
library(stringr)            # string manipulation
library(readxl)             # reading Excel files
library(cld2)               # language detection
library(DAMisc)             # summary statistics helpers
library(broom)              # tidy model output
library(patchwork)          # combining ggplots
library(stargazer)          # regression tables

```

### Load Data

```{r load-data, warning = FALSE}

# Set working directory to project root
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("..")

# Load the final cleaned dataset (produced by cleaning.Rmd)
# Contains: final_data, dnk, nld_cong, ger, esp, gbr, swe, nld_parl, eus
load("data/cleaned/cole_final_cleaned_data.RData")

```

---

## Descriptive Statistics

These tables match the appendix tables in the original paper. Each table shows
summary statistics for the five key variables across all 8 corpora.

```{r descriptive-stats, warning = FALSE}

# Function to produce a summary statistics table for a given dataset
sum_stats_table <- function(df, title) {
  vars   <- c("flesch_kincaid", "left_right_scale", "lib_con_scale",
               "speech_year", "in_gov")
  labels <- c("Complexity", "Left-Right", "Progressive-Conservative",
               "Year", "In government")

  map2_dfr(vars, labels, ~ {
    x <- df[[.x]]
    tibble(
      Statistic  = .y,
      N          = format(sum(!is.na(x)), big.mark = ","),
      Mean       = round(mean(x, na.rm = TRUE), 3),
      `St. Dev.` = round(sd(x, na.rm = TRUE), 3),
      Min        = round(min(x, na.rm = TRUE), 3),
      `Pctl(25)` = round(quantile(x, .25, na.rm = TRUE), 3),
      `Pctl(75)` = round(quantile(x, .75, na.rm = TRUE), 3),
      Max        = round(max(x, na.rm = TRUE), 3)
    )
  }) |>
    kable(caption = title, align = c("l", rep("r", 7)))
}

# Generate all 8 descriptive statistics tables
dnk      |> sum_stats_table("Table A.1: Party congresses DNK")
nld_cong |> sum_stats_table("Table A.2: Party congresses NLD")
ger      |> sum_stats_table("Table A.3: Parliament GER")
esp      |> sum_stats_table("Table A.4: Parliament ESP")
gbr      |> sum_stats_table("Table A.5: Parliament GBR")
swe      |> sum_stats_table("Table A.6: Parliament SWE")
nld_parl |> sum_stats_table("Table A.7: Parliament NLD")
eus      |> sum_stats_table("Table A.8: Heads of Government")

```

---

## Figure 1: Descriptive Information

This figure shows mean complexity scores for selected EU speakers, illustrating
the the Flesch-Kincaid score's ability to measure complexity.


```{r figure-1, fig.keep="none"}

fig_1_rep <- eus |>
  filter(speaker %in% c("G. Brown", "D. Cameron",
                         "J.L.R. Zapatero", "M. Rajoy")) |>
  group_by(speaker) |>
  summarise(
    mean_complexity = mean(flesch_kincaid_unstand, na.rm = TRUE),
    se = sd(flesch_kincaid, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  mutate(
    speaker = fct_reorder(speaker, mean_complexity),
    ideology = case_when(
      speaker %in% c("G. Brown", "J.L.R. Zapatero") ~ "Liberal",
      speaker %in% c("D. Cameron", "M. Rajoy")       ~ "Conservative"
    )
  ) |>
  ggplot(aes(x = mean_complexity, y = speaker, fill = ideology)) +
  geom_col() +
  geom_errorbar(aes(xmin = mean_complexity - 1.96 * se,
                    xmax = mean_complexity + 1.96 * se),
                width = 0.2) +
  scale_fill_manual(
    values = c("Liberal" = "#E69F00", "Conservative" = "#56B4E9"),
    name = NULL
  ) +
  labs(x = "Mean Complexity", y = NULL,
       title = "Fig 1: Descriptive information on linguistic complexity") +
  theme_minimal() +
  theme(legend.position = "bottom")

fig_1_rep

# Save figure
ggsave("../figures/cole-fig-1.png", fig_1_rep, width = 8, height = 5)

```

---

## Figure 2: Effect of Ideology on Complexity

This replicates the paper's main result figure. For each of the 8 corpora, we
run an OLS regression of standardized FK scores on standardized ideology
(lib-con and left-right), year, and governing party status, then plot the
coefficients with 95% confidence intervals.

```{r figure-2, fig.width = 12, fig.height = 8, fig.keep="none"}

# Define all 8 datasets
datasets <- list(
  "Parliament GER"         = ger,
  "Parliament ESP"         = esp,
  "Parliament GBR"         = gbr,
  "Parliament SWE"         = swe,
  "Parliament NLD"         = nld_parl,
  "Party congresses DNK"   = dnk,
  "Party congresses NLD"   = nld_cong,
  "Heads of Government"    = eus
)

# Run OLS for each dataset, extract ideology coefficients
results <- map_dfr(names(datasets), function(name) {
  df <- datasets[[name]] |>
    mutate(across(c(flesch_kincaid, lib_con_scale,
                    left_right_scale, speech_year), scale))

  model <- lm(flesch_kincaid ~ lib_con_scale + left_right_scale +
               speech_year + in_gov, data = df)

  tidy(model, conf.int = TRUE) |>
    filter(term %in% c("lib_con_scale", "left_right_scale")) |>
    mutate(dataset = name)
})

# Add panel grouping for faceting (Parliament vs Congress/HoG)
results <- results |>
  mutate(
    panel = ifelse(str_detect(dataset, "Parliament"),
                   "Parliament", "Congress and Heads of Government"),
    panel = factor(panel,
                   levels = c("Parliament", "Congress and Heads of Government")),
    dataset = factor(dataset, levels = rev(c(
      "Parliament GER", "Parliament ESP", "Parliament GBR",
      "Parliament SWE", "Parliament NLD",
      "Party congresses DNK", "Party congresses NLD", "Heads of Government"
    )))
  )

# Panel a: Liberal-Conservative ideology effect
p2a <- results |>
  filter(term == "lib_con_scale") |>
  ggplot(aes(x = estimate, y = dataset)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~panel, scales = "free_y", ncol = 1) +
  labs(x = "Effect of lib-cons ideology", y = NULL, title = "a") +
  theme_minimal()

# Panel b: Left-Right ideology effect
p2b <- results |>
  filter(term == "left_right_scale") |>
  ggplot(aes(x = estimate, y = dataset)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~panel, scales = "free_y", ncol = 1) +
  labs(x = "Effect of left-right ideology", y = NULL, title = "b") +
  theme_minimal()

fig_2_plot <- p2a + p2b
fig_2_plot

ggsave("../figures/cole-fig-2.png", fig_2_plot, width = 12, height = 8)

```

---

## Figure 3: Complexity Over Time

This figure shows smoothed trends in speech complexity over time, separately
for parliamentary speeches and party congress speeches.

```{r figure-3, fig.width = 12, fig.height = 6, fig.keep="none"}

# Panel a: Parliamentary speeches (5 countries)
p3a <- bind_rows(
  esp      |> mutate(corpus = "ESP"),
  gbr      |> mutate(corpus = "GBR"),
  ger      |> mutate(corpus = "GER"),
  nld_parl |> mutate(corpus = "NLD"),
  swe      |> mutate(corpus = "SWE")
) |>
  mutate(
    corpus = factor(corpus, levels = c("ESP", "GBR", "GER", "NLD", "SWE")),
    year = speech_year + 1988  # convert zero-based year back to calendar year
  ) |>
  ggplot(aes(x = year, y = flesch_kincaid_unstand)) +
  geom_smooth(method = "loess", se = FALSE, color = "red", span = 0.5) +
  facet_wrap(~corpus, nrow = 1, scales = "free_x") +
  scale_x_continuous(breaks = c(1990, 2000, 2010),
                     labels = c("'90", "'00", "'10")) +
  labs(x = NULL, y = "Complexity", title = "a") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())

# Panel b: Party congress speeches (Denmark and Netherlands)
p3b <- bind_rows(
  dnk      |> mutate(corpus = "DNK"),
  nld_cong |> mutate(corpus = "NLD")
) |>
  mutate(
    corpus = factor(corpus, levels = c("DNK", "NLD")),
    year = speech_year + 1944  # convert zero-based year back to calendar year
  ) |>
  ggplot(aes(x = year, y = flesch_kincaid_unstand)) +
  geom_smooth(method = "loess", se = FALSE, color = "red", span = 0.5) +
  facet_wrap(~corpus, nrow = 1, scales = "free_x") +
  scale_x_continuous(breaks = c(1960, 1980, 2000),
                     labels = c("'60", "'80", "'00")) +
  labs(x = NULL, y = "Complexity", title = "b") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())

fig_3_plot <- p3a + p3b
fig_3_plot

ggsave("../figures/cole-fig-3.png", fig_3_plot, width = 12, height = 6)

```

---

## Figure 4: Party Fixed Effects

Panel (a) shows OLS coefficients for the effect of conservatism with party
fixed effects included. Panel (b) shows within-party loess trends for ideology
and complexity over time for two illustrative parties.

```{r figure-4, fig.width = 12, fig.height = 6, fig.keep="none"}

# --- Panel a: Party fixed effects (excluding EUSpeech) ---
datasets_noeus <- list(
  "Parliament GER"       = ger,
  "Parliament ESP"       = esp,
  "Parliament GBR"       = gbr,
  "Parliament SWE"       = swe,
  "Parliament NLD"       = nld_parl,
  "Party congresses DNK" = dnk,
  "Party congresses NLD" = nld_cong
)

# Run OLS with party FE for each corpus
results_fe <- map_dfr(names(datasets_noeus), function(name) {
  df <- datasets_noeus[[name]] |>
    mutate(across(c(flesch_kincaid, lib_con_scale,
                    left_right_scale, speech_year), scale))

  model <- lm(flesch_kincaid ~ lib_con_scale + left_right_scale +
               speech_year + in_gov + factor(party), data = df)

  tidy(model, conf.int = TRUE) |>
    filter(term == "lib_con_scale") |>
    mutate(dataset = name)
})

results_fe <- results_fe |>
  mutate(
    panel = ifelse(str_detect(dataset, "Parliament"), "Parliament", "Congress"),
    panel = factor(panel, levels = c("Parliament", "Congress")),
    dataset = factor(dataset, levels = rev(c(
      "Parliament GER", "Parliament ESP", "Parliament GBR",
      "Parliament SWE", "Parliament NLD",
      "Party congresses DNK", "Party congresses NLD"
    )))
  )

p4a <- results_fe |>
  ggplot(aes(x = estimate, y = dataset)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "solid") +
  facet_wrap(~panel, scales = "free_y", ncol = 1) +
  labs(x = "Fixed effect of conservatism", y = NULL, title = "a") +
  theme_minimal()

# --- Panel b: Within-party trends (SD in Denmark, VVD in Netherlands) ---
sd_data <- dnk |>
  filter(party == "SD") |>
  mutate(
    year       = speech_year + 1944,
    complexity = scale(flesch_kincaid)[, 1],
    ideology   = scale(lib_con_scale)[, 1]
  ) |>
  select(year, complexity, ideology) |>
  pivot_longer(c(complexity, ideology),
               names_to = "variable", values_to = "value") |>
  mutate(corpus = "Social Democratic Party\n(DK)")

vvd_data <- nld_cong |>
  filter(party == "VVD") |>
  mutate(
    year       = speech_year + 1944,
    complexity = scale(flesch_kincaid)[, 1],
    ideology   = scale(lib_con_scale)[, 1]
  ) |>
  select(year, complexity, ideology) |>
  pivot_longer(c(complexity, ideology),
               names_to = "variable", values_to = "value") |>
  mutate(corpus = "Liberal Party\n(NL)")

p4b <- bind_rows(sd_data, vvd_data) |>
  mutate(corpus = factor(corpus,
    levels = c("Social Democratic Party\n(DK)", "Liberal Party\n(NL)")
  )) |>
  ggplot(aes(x = year, y = value, color = variable)) +
  geom_smooth(method = "loess", se = FALSE, span = 0.5) +
  facet_wrap(~corpus, ncol = 1) +
  scale_color_manual(values = c("complexity" = "blue", "ideology" = "red")) +
  labs(x = NULL, y = "Standardized values", color = NULL, title = "b") +
  theme_minimal() +
  theme(legend.position = "bottom",
        strip.text.y = element_text(angle = 0))

fig_4_plot <- p4a + p4b
fig_4_plot

ggsave("../figures/cole-fig-4.png", fig_4_plot, width = 12, height = 6)

```

---

## OLS Regression Tables

### Standard OLS (No Fixed Effects)

These models regress standardized Flesch-Kincaid scores on ideology (left-right
and lib-con), year, and governing party status. The EUSpeech model excludes
year and governing status due to the smaller sample of heads of government.

```{r ols-table-1, results = 'asis'}

m_dnk      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = dnk)
m_nld_cong <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = nld_cong)
m_ger      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = ger)
m_esp      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = esp)
m_gbr      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = gbr)
m_swe      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = swe)
m_nld_parl <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                  speech_year + in_gov, data = nld_parl)
m_eus      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale,
                  data = eus)

stargazer(m_dnk, m_nld_cong, m_ger, m_esp,
          title = "Table C.1: OLS regression results",
          column.labels = c("DNK Congress", "NLD Congress",
                            "GER Parl", "ESP Parl"),
          covariate.labels = c("Left-Right", "Progressive-Conservative",
                               "Year", "In government"),
          type = "html",
          star.cutoffs = c(0.1, 0.05, 0.01))

stargazer(m_gbr, m_swe, m_nld_parl, m_eus,
          title = "Table C.1 continued",
          column.labels = c("GBR Parl", "SWE Parl", "NLD Parl", "PM"),
          covariate.labels = c("Left-Right", "Progressive-Conservative",
                               "Year", "In government"),
          type = "html",
          star.cutoffs = c(0.1, 0.05, 0.01))

```

### Speaker Fixed Effects

These models add speaker fixed effects to control for individual speaking
styles. This tests whether the ideology-complexity relationship holds
*within speakers* as their party's ideology shifts.

```{r ols-table-2, results = 'asis'}

m_dnk_fe      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = dnk)
m_nld_cong_fe <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = nld_cong)
m_ger_fe      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = ger)
m_esp_fe      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = esp)
m_gbr_fe      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = gbr)
m_swe_fe      <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = swe)
m_nld_parl_fe <- lm(flesch_kincaid ~ left_right_scale + lib_con_scale +
                     speech_year + in_gov + factor(speaker), data = nld_parl)

stargazer(m_dnk_fe, m_nld_cong_fe, m_ger_fe, m_esp_fe,
          title = "Table D.1: Speaker fixed effects regression results",
          column.labels = c("DNK Congress", "NLD Congress",
                            "GER Parl", "ESP Parl"),
          covariate.labels = c("Left-Right", "Progressive-Conservative",
                               "Year", "In government"),
          omit = "factor\\(speaker\\)",
          add.lines = list(c("Speaker Fixed Effects?",
                             "Yes", "Yes", "Yes", "Yes")),
          type = "html",
          star.cutoffs = c(0.1, 0.05, 0.01))

stargazer(m_gbr_fe, m_swe_fe, m_nld_parl_fe,
          title = "Table D.1 continued",
          column.labels = c("GBR Parl", "SWE Parl", "NLD Parl"),
          covariate.labels = c("Left-Right", "Progressive-Conservative",
                               "Year", "In government"),
          omit = "factor\\(speaker\\)",
          add.lines = list(c("Speaker Fixed Effects?",
                             "Yes", "Yes", "Yes")),
          type = "html",
          star.cutoffs = c(0.1, 0.05, 0.01))

```
