---
title: "Data Cleaning – Replication of Schoonvelde et al. (2019)"
author: "Colton Stevenson"
date: "April 3rd 2026"
output:
  html_document:
    theme: darkly
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: true
    toc_depth: 5
    highlight: zenburn
---

## Overview

This notebook handles all data cleaning for the replication of:

> Schoonvelde, M., Brosius, A., &"; van Atteveldt, W. (2019). Liberals lecture,
> conservatives communicate: Analyzing complexity and ideology in 381,609
> political speeches. *PLOS ONE*, 14(2), e0208450.

We start from publicly available raw datasets, clean them to match the paper's
specifications, calculate ideology scores from the Manifesto Project, compute
Flesch-Kincaid complexity scores, and produce a final analysis-ready `.RData`
file.

**Inputs:**

- `data/raw/ParlSpeech/` — Five European parliament corpora
- `data/raw/EUSpeech/supercorpus_unstemmed_V4.RData` — EU elite speeches (unstemmed V4)
- `data/raw/PartyCongressSpeech/corpus_DK_NL.Rdata` — Danish and Dutch party congress speeches
- `data/raw/ManifestoProject/ManifestoProject2019.xlsx` — Manifesto Project (2019 version)

**Output:**

- `data/cleaned/cole_final_cleaned_data.RData` — Final cleaned dataset with all variables

---

## Setup

### Libraries

```{r load-libraries, message = FALSE, warning = FALSE}

library(tidyverse)          # data wrangling
library(tidylog)            # verbose output for dplyr operations
library(tidytext)           # tidy text manipulation
library(textdata)           # text datasets
library(quanteda)           # corpus and tokenization
library(quanteda.textstats) # readability scores (Flesch-Kincaid)
library(wesanderson)        # colour palettes
library(knitr)              # table formatting
library(stringr)            # string manipulation
library(readxl)             # reading .xlsx files
library(cld2)               # language detection

```

### Load Raw Data

All paths are relative to the project root.

```{r load-data, warning = FALSE}

# Set working directory to the project root (one level up from /code)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("..")

# ---------------------------------------------------------------------------
# ParlSpeech: Speeches from five European parliaments
# Source: https://doi.org/10.7910/DVN/E4RSP9
# ---------------------------------------------------------------------------
load("data/raw/ParlSpeech/Corp_Bundestag.Rdata")      # German Bundestag
load("data/raw/ParlSpeech/Corp_Congresso.Rdata")       # Spanish Congresso
load("data/raw/ParlSpeech/Corp_HouseOfCommons.Rdata")  # British House of Commons
load("data/raw/ParlSpeech/Corp_Riksdag.Rdata")         # Swedish Riksdag
load("data/raw/ParlSpeech/Corp_TweedeKamer.Rdata")     # Dutch Tweede Kamer

# ---------------------------------------------------------------------------
# EUSpeech: Speeches from EU elites and heads of state/government
# Source: https://doi.org/10.7910/DVN/GKABNU (updated V4 with unstemmed text)
# Note: The DOI cited in the paper (DVN/XPCVEI) links to the 2016 version,
#       which only contains stemmed/cleaned text. The V4 unstemmed corpus is
#       the correct one to use for Flesch-Kincaid scoring.
# ---------------------------------------------------------------------------
load("data/raw/EUSpeech/supercorpus_unstemmed_V4.RData")

# ---------------------------------------------------------------------------
# Party Congress Speeches: Denmark and the Netherlands (1945-2017)
# Source: https://doi.org/10.7910/DVN/SFZI8O
# ---------------------------------------------------------------------------
load("data/raw/PartyCongressSpeech/corpus_DK_NL.Rdata")

# ---------------------------------------------------------------------------
# Manifesto Project: Party ideology scores (2019 version)
# Source: https://manifesto-project.wzb.eu/ (registration required)
# ---------------------------------------------------------------------------
manifesto <- read_excel("data/raw/ManifestoProject/ManifestoProject2019.xlsx")

```

---

## Pre-Cleaning

Before the substantive cleaning, we simplify each dataset by selecting only the
variables we need and merging the five ParlSpeech corpora into one dataframe.
We also set up optional sampling for faster testing during development.

```{r pre-cleaning}

# Combine all five ParlSpeech corpora into one dataframe
parlspeech_raw <- bt.corpus |>
  rbind(cong.corpus, hoc.corpus, rd.corpus, tk.corpus) |>
  select(date, speaker, party, chair, terms, text, parliament)

# EUSpeech: extract the documents dataframe from the corpus list object
euspeech_raw <- corpus_unstemmed[["documents"]]

# Party Congresses: extract documents and keep key variables
partycong_raw <- corpus[["documents"]] |>
  select(year, speaker, party, texts, country)

# Manifesto Project: keep country, date, party info, and all "per" issue codes
manifesto_raw <- manifesto |>
  select(countryname, date, edate, partyname, partyabbrev, starts_with("per"))

```

### Sampling (For Testing)

During development, we used a sample of 10,000 speeches per dataset to speed up
testing. For the final results, we use the full dataset (set `n_sam <- Inf`).

```{r sampling}

# Toggle between sample and full data:
 n_sam <- 10000  # sample for testing
# n_sam <- Inf      # full dataset for final results (ParlSpeech takes ~7 min)

parlspeech_sample <- slice_sample(parlspeech_raw, n = n_sam)
euspeech_sample   <- slice_sample(euspeech_raw,   n = n_sam)
partycong_sample  <- slice_sample(partycong_raw,   n = n_sam)

```

---

## Cleaning

### ParlSpeech

The paper specifies two filters for parliamentary speeches:

1. Exclude all chairperson speeches (they organize debate, not substantive)
2. Keep only speeches with at least 10 sentences of at least 5 words

We also remap party abbreviations to match the Manifesto Project naming
conventions, which is necessary for the ideology merge later.

```{r clean-parlspeech}

parlspeech_clean <- parlspeech_sample |>
  filter(chair == "FALSE") |> # Removing Chair Speeches
  # Regex: match sentences with >= 5 words ending in punctuation; keep if >= 10
  filter(str_count(text, "\\b\\w+\\b(?:\\s+\\b\\w+\\b){4,}\\s*[.!?;:]") >= 10) |>
  mutate(
    # Add country names for later merging with Manifesto data
    countryname = case_when(
      parliament == "NL-TweedeKamer" ~ "Netherlands",
      parliament == "ES-Congresso"   ~ "Spain",
      parliament == "UK-HoC"         ~ "United Kingdom",
      parliament == "SE-Riksdag"     ~ "Sweden",
      parliament == "DE-Bundestag"   ~ "Germany"
    )
  ) |>
  mutate(
    # Remap party abbreviations to match Manifesto Project conventions
    speech_year = year(as.Date(date)),
    party = case_when(
      # United Kingdom
      party == "Con"        ~ "Conservatives",
      party == "Lab"        ~ "Labour",
      party == "LibDem"     ~ "LibDems",
      party == "PlaidCymru" ~ "PC",
      # Germany — PDS/LINKE renamed to LINKE after 2007
      party == "GRUENE"     ~ "90/Greens",
      party == "PDS/LINKE" & speech_year >= 2007 ~ "LINKE",
      party == "PDS/LINKE" & speech_year < 2007  ~ "PDS",
      # Sweden
      party == "S"   ~ "SAP",
      party == "C"   ~ "CP",
      party == "KD"  ~ "Kd",
      party == "M"   ~ "MSP",
      party == "NYD" ~ "NyD",
      party == "L"   ~ "L",
      # Netherlands
      party == "D66" ~ "D'66",
      # Spain
      party == "PNV"    ~ "PNV/EAJ",
      party == "IU-ICV" ~ "IU",
      TRUE ~ party
    )
  )

```

### EUSpeech

The EUSpeech dataset required more work than ParlSpeech because:

- The raw data does not include party or full speaker information
- The DOI in the paper links to a 2016 version with stemmed/cleaned text
- We need the unstemmed V4 version for accurate Flesch-Kincaid scoring

#### Adding Speaker Information

The original dataset only includes speaker names (often abbreviated) without
party affiliations. We built a lookup table mapping each speaker to their party
and Manifesto Project abbreviation.

**Note: An LLM (Claude) was used to assist with researching the party
affiliations and abbreviations for each speaker in the dataset. All entries
were verified against publicly available sources.**

```{r euspeech-speaker-info}

# Lookup table: speaker -> party name and Manifesto Project abbreviation
# Built with LLM assistance for researching ~60 speakers' party affiliations
speaker_info <- tribble(
  ~speaker,               ~party,                                               ~partyabbrev,
  "J.L.R. Zapatero",     "Spanish Socialist Workers' Party",                   "PSOE",
  "M. Rutte",            "People's Party for Freedom and Democracy",           "VVD",
  "Francesco Speroni",   "Northern League",                                    "LN",
  "A. Merkel",           "Christian Democratic Union/Christian Social Union",  "CDU/CSU",
  "Marine Le Pen",       "National Front",                                     "FN",
  "European Commission", "EU Institution",                                     NA,
  "F. Hollande",         "Socialist Party",                                    "PS",
  "Michal Kaminski",     "Law and Justice",                                    "PiS",
  "Martin Schulz",       "Social Democratic Party of Germany",                 "SPD",
  "M. Rajoy",            "People's Party",                                     "PP",
  "Francis Wurtz",       "French Communist Party",                             "PCF",
  "IMF",                 "International Institution",                          NA,
  "M. Monti",            "Civic Choice",                                       "SC",
  "Rebecca Harms",       "Alliance'90/Greens",                                 "90/Greens",
  "G.A. Papandreou",     "Panhellenic Socialist Movement",                     "PASOK",
  "H. van Rompuy",       "Christian Democratic and Flemish",                   "CD&V",
  "S. Berlusconi",       "People of Freedom",                                  "PdL",
  "G. Brown",            "Labour Party",                                       "Labour",
  "J.P. Balkenende",     "Christian Democratic Appeal",                        "CDA",
  "N. Sarkozy",          "Union for a Popular Movement",                       "UMP",
  "Guy Verhofstadt",     "Open Flemish Liberals and Democrats",                "openVLD",
  "A. Samaras",          "New Democracy",                                      "ND",
  "R. Prodi",            "Democratic Party",                                   "PD",
  "Monica Frassoni",     "Alliance'90/Greens",                                 "90/Greens",
  "Daniel Cohn-Bendit",  "Alliance'90/Greens",                                 "90/Greens",
  "ECB",                 "EU Institution",                                     NA,
  "A. Tsipras",          "Coalition of the Radical Left",                      "SYRIZA",
  "Martin Callanan",     "Conservative Party",                                 "Conservatives",
  "Cristiana Muscardini","National Alliance",                                  "AN",
  "B. Sobotka",          "Czech Social Democratic Party",                      "CSSD",
  "Passos Coelho",       "Social Democratic Party",                            "PSD",
  "D. Cameron",          "Conservative Party",                                 "Conservatives",
  "Manfred Weber",       "Christian Democratic Union/Christian Social Union",  "CDU/CSU",
  "Brian Crowley",       "Soldiers of Destiny",                                NA,
  "Jan Zahradil",        "Civic Democratic Party",                             "ODS",
  "Hannes Swoboda",      "Austrian Social Democratic Party",                   "SPO",
  "Graham Watson",       "Liberal Democrats",                                  "LibDems",
  "Joseph Daul",         "Union for a Popular Movement",                       "UMP",
  "Gabriele Zimmer",     "The Left",                                           "LINKE",
  "David Borrelli",      "Five Star Movement",                                 "M5S",
  "Nigel Farage",        "United Kingdom Independence Party",                  "UKIP",
  "L. Papademos",        "Independent",                                        NA,
  "Jose Socrates",       "Socialist Party",                                    "PS",
  "Lothar Bisky",        "The Left",                                           "LINKE",
  "Kathy Sinnott",       "Independent",                                        NA,
  "Marcel de Graaff",    "Party of Freedom",                                   "PVV",
  "Philippe Lamberts",   "Ecologists",                                         "ECOLO",
  "E. Letta",            "Democratic Party",                                   "PD",
  "Donald Tusk",         "Civic Platform",                                     "PO",
  "Gianni Pittella",     "Democratic Party",                                   "PD",
  "D. Tusk",             "Civic Platform",                                     "PO",
  "J. Fischer",          "Alliance'90/Greens",                                 "90/Greens",
  "T. Blair",            "Labour Party",                                       "Labour",
  "M. Renzi",            "Democratic Party",                                   "PD",
  "P. Necas",            "Civic Democratic Party",                             "ODS",
  "Syed Kamall",         "Conservative Party",                                 "Conservatives",
  "J. Chirac",           "Rally for the Republic",                             "RPR",
  "Antonia Costa",       "Socialist Party",                                    "PS",
  "Beaty Szydlo",        "Law and Justice",                                    "PiS",
  "Ewa Kopacz",          "Civic Platform",                                     "PO"
)

# Join speaker info, recode country, filter for English non-translated speeches
euspeech_clean <- euspeech_sample |>
  left_join(speaker_info, by = "speaker") |>
  mutate(country = recode(country, "Great Britain" = "United Kingdom")) |>
  filter(language == "en" & translation == 0) |>    # filtering language + no translation
  filter(country %in% c(   # Setting right countries
    "Czech Republic", "Germany", "Greece", "Netherlands", "Italy",
    "Spain", "United Kingdom", "Poland", "Portugal", "France"
  ))

```

#### Filtering Speakers

The paper excludes Italian PM Prodi (only 3 English speeches) and Portuguese PM
Passos Coelho (only 6 speeches) due to insufficient data.

```{r euspeech-filtering}

euspeech_clean <- euspeech_clean |>
  filter(speaker != "Passos Coelho") |>
  filter(speaker != "R. Prodi")

```

### Party Congresses

We filter to only the parties specified in the paper and handle the historical
merger of ARP, CHU, and KVP into the Christian Democratic Appeal (CDA). We keep
a separate `manifesto_party` column so that pre-CDA speeches (before ~1980) can
be matched to the correct historical party's ideology scores.

```{r clean-party-congresses}

partycong_clean <- partycong_sample |>
  # Keep only the parties specified in the paper
  filter(party %in% c(
    "DF", "EL", "S", "SF", "V",                             # Denmark
    "SP", "Groenlinks", "PvdA", "VVD", "ARP", "CHU", "KVP", # Netherlands
    "CDA", "D66", "PVV"
  )) |>
  mutate(
    # Preserve original abbreviation for Manifesto matching (pre-CDA parties)
    manifesto_party = case_when(
      party == "ARP" ~ "ARP",
      party == "CHU" ~ "CHU",
      party == "KVP" ~ "KVP",
      party == "CDA" ~ "CDA",
      TRUE           ~ party
    ),
    # Merge ARP/CHU/KVP into CDA for the analysis (as per the paper)
    party = case_when(
      party %in% c("ARP", "CHU", "KVP", "CDA") ~ "CDA",
      TRUE ~ party
    )
  ) |>
  mutate(
    # Add full party names for readability
    party_name = case_when(
      party == "DF"         ~ "Danish People's Party",
      party == "EL"         ~ "Unity List",
      party == "S"          ~ "Social Democrats",
      party == "SF"         ~ "Socialist People's Party",
      party == "V"          ~ "Venstre",
      party == "SP"         ~ "Socialist Party",
      party == "Groenlinks" ~ "Green Left",
      party == "PvdA"       ~ "Labour Party",
      party == "VVD"        ~ "VVD",
      party == "CDA"        ~ "Christian Democratic Appeal",
      party == "D66"        ~ "D66",
      party == "PVV"        ~ "Freedom Party"
    )
  ) |>
  mutate(
    countryname = case_when(
      country == "DK" ~ "Denmark",
      country == "NL" ~ "Netherlands"
    )
  ) |>
  mutate(
    # Remap abbreviations to match Manifesto Project conventions
    party = case_when(
      party == "S"          & countryname == "Denmark"     ~ "SD",
      party == "D66"        & countryname == "Netherlands" ~ "D'66",
      party == "Groenlinks" & countryname == "Netherlands" ~ "GL",
      TRUE ~ party
    ),
    manifesto_party = case_when(
      manifesto_party == "S"          & countryname == "Denmark"     ~ "SD",
      manifesto_party == "D66"        & countryname == "Netherlands" ~ "D'66",
      manifesto_party == "Groenlinks" & countryname == "Netherlands" ~ "GL",
      TRUE ~ manifesto_party
    )
  )

```

---

## Ideology Scores (Manifesto Project)

The paper uses two ideological dimensions derived from the Manifesto Project:

The Liberal Conservative scale is calculated using this formula:

$$ideo=log(\sum{conservative~issues + 0.5)} - log(\sum{libeal~issues + 0.5)}$$

In which conservative and liberal issues are decided as such:

- **Conservative issues:** 
	- anti-EU
	- anti-immigration
	- pro-national way of life
	- pro-traditional morality
	- anti-multiculturalism
	- pro-military
	- anti-internationalism
	- pro-Freedom and Human Right
	- pro-political authority
	- pro-law and order
- **Liberal Issues:**
	- pro-EU
	- pro-immigration
	- anti-national way of life
	- anti-traditional morality
	- pro-multiculturalism
	- antimilitary
	- pro-internationalism
	- anti-imperialism
	- pro-peace
	- pro-environment
	- pro-culture
	- support for under-privileged minority groups.
	
The economic left / right scale is calculated using this formula:

$$ideo=log(\sum{left~issues + 0.5)} - log(\sum{right~issues + 0.5)}$$

In which right and left economic issues are decided as such:

- **Right issues:** 
	- free-market economy
	- incentives
	- protectionism: negative
	- economic growth: positive
	- economic orthodoxy
	- welfare state limitation
	- labour groups: negative
- **Left Issues:**
	- market regulation
	- economic planning
	- corporatism
	- protectionism: positive
	- Keynesian demand management
	- controlled economy
	- nationalisation
	- Marxist analysis
	- welfare state expansion
	- education expansion
	- support for labour groups

```{r ideology-scores}

manifesto_clean <- manifesto_raw |>
  # Normalize the apostrophe in D'66
  mutate(partyabbrev = str_replace_all(partyabbrev, "\u2019", "'")) |>
  mutate(
    # --- Liberal-Conservative cultural dimension ---
    # Conservative issues (Manifesto Project codes)
    conservative_sum = per110 + per601 + per603 + per608 + per104 +
                       per109 + per201 + per305 + per605,
    # Liberal issues
    liberal_sum      = per108 + per602 + per604 + per607 + per105 +
                       per107 + per103 + per106 + per501 + per502 + per705,
    lib_con_scale    = log(conservative_sum + 0.5) - log(liberal_sum + 0.5),

    # --- Left-Right economic dimension ---
    right_sum = per401 + per402 + per407 + per410 + per414 + per505 + per702,
    left_sum  = per403 + per404 + per405 + per406 + per409 + per412 +
                per413 + per415 + per504 + per506 + per701,
    left_right_scale = log(left_sum + 0.5) - log(right_sum + 0.5)
  ) |>
  select(countryname, date, edate, partyname, partyabbrev,
         lib_con_scale, left_right_scale)

```

### Merging Ideology Into Speech Datasets

Ideology scores change over time (updated each election), so we match each
speech to the most recent manifesto published *before* the speech date. This
is done by joining on party abbreviation and country, then filtering for the
latest manifesto that predates the speech.

```{r merge-ideology}

# Prepare manifesto dates for time-aware joining
manifesto_clean <- manifesto_clean |>
  mutate(
    manifesto_year  = as.integer(str_sub(date, 1, 4)),
    manifesto_month = as.integer(str_sub(date, 5, 6))
  ) |>
  rename(manifesto_date = date)

# --- ParlSpeech ---
parlspeech_ideology <- parlspeech_clean |>
  mutate(
    speech_year  = year(as.Date(date)),
    speech_month = month(as.Date(date))
  ) |>
  left_join(manifesto_clean, by = c("party" = "partyabbrev", "countryname")) |>
  # Keep only manifesto scores from before or during the speech date
  filter(
    manifesto_year < speech_year |
    (manifesto_year == speech_year & manifesto_month <= speech_month)
  ) |>
  # Keep the most recent manifesto for each speech
  group_by(speaker, date) |>
  slice_max(order_by = manifesto_year * 100 + manifesto_month) |>
  ungroup()

# --- Party Congresses ---
# Uses manifesto_party (not party) so pre-CDA parties get their own scores
partycong_ideology <- partycong_clean |>
  mutate(speech_year = as.integer(year)) |>
  left_join(manifesto_clean,
            by = c("manifesto_party" = "partyabbrev", "countryname")) |>
  filter(manifesto_year <= speech_year) |>
  group_by(speaker, year) |>
  slice_max(order_by = manifesto_year * 100 + manifesto_month) |>
  ungroup()

# --- EUSpeech ---
euspeech_ideology <- euspeech_clean |>
  mutate(
    date_parsed  = as.Date(date, format = "%d-%m-%Y"),
    speech_year  = year(date_parsed),
    speech_month = month(date_parsed)
  ) |>
  left_join(manifesto_clean,
            by = c("partyabbrev", "country" = "countryname")) |>
  filter(
    manifesto_year < speech_year |
    (manifesto_year == speech_year & manifesto_month <= speech_month)
  ) |>
  group_by(speaker, date) |>
  slice_max(order_by = manifesto_year * 100 + manifesto_month) |>
  ungroup()

```

---

## Complexity (Flesch-Kincaid Grade Level)

Speech complexity is measured using the Flesch-Kincaid Grade Level formula:

$$score = 0.39 \times \frac{\text{total words}}{\text{total sentences}} + 11.8 \times \frac{\text{total syllables}}{\text{total words}} - 15.59$$

We use `quanteda.textstats::textstat_readability()` for this calculation.

**Note:** ParlSpeech is very large (~700k speeches after filtering), so we split
it into 3 chunks to avoid exceeding 32 GB of RAM.

```{r complexity, message = FALSE}

# --- ParlSpeech: chunked to avoid memory issues ---
n <- nrow(parlspeech_ideology)
chunk_size <- ceiling(n / 3)
chunks <- split(parlspeech_ideology, ceiling(seq_len(n) / chunk_size))

fk_scores <- lapply(seq_along(chunks), function(i) {
  message("Processing ParlSpeech chunk ", i, " of ", length(chunks))
  corp <- corpus(chunks[[i]], text_field = "text")
  textstat_readability(corp, measure = "Flesch.Kincaid")
}) |> bind_rows()

parlspeech_ideology$flesch_kincaid <- fk_scores$Flesch.Kincaid

# --- EUSpeech ---
corpus_eu <- corpus(euspeech_ideology, text_field = "texts")
fk_eu     <- textstat_readability(corpus_eu, measure = "Flesch.Kincaid")
euspeech_ideology$flesch_kincaid <- fk_eu$Flesch.Kincaid

# --- Party Congresses ---
corpus_pc <- corpus(partycong_ideology, text_field = "texts")
fk_pc     <- textstat_readability(corpus_pc, measure = "Flesch.Kincaid")
partycong_ideology$flesch_kincaid <- fk_pc$Flesch.Kincaid

# Quick sanity check: means should be in the 10-15 range
message("ParlSpeech mean FK: ", round(mean(parlspeech_ideology$flesch_kincaid), 2))
message("EUSpeech mean FK: ",   round(mean(euspeech_ideology$flesch_kincaid), 2))
message("PartyCong mean FK: ",  round(mean(partycong_ideology$flesch_kincaid), 2))

```

---

## Governing Party Dummy

We construct a lookup table of which parties were in government during each
time period, then flag speeches where the speaker's party was in power.

**Note: An LLM (Claude) was used to assist with compiling the governing
coalition lookup table. All entries were verified against publicly available
sources (government archives, Wikipedia).**

```{r governing-party}

gov_periods <- tribble(
  ~countryname,    ~party,          ~gov_start,             ~gov_end,

  # --- GERMANY ---
  # Kohl (CDU/CSU + FDP): 1990-1998
  "Germany",       "CDU/CSU",   as.Date("1990-01-01"), as.Date("1998-10-26"),
  "Germany",       "FDP",       as.Date("1990-01-01"), as.Date("1998-10-26"),
  # Schroeder (SPD + Greens): 1998-2005
  "Germany",       "SPD",       as.Date("1998-10-27"), as.Date("2005-11-21"),
  "Germany",       "90/Greens", as.Date("1998-10-27"), as.Date("2005-11-21"),
  # Merkel I (CDU/CSU + SPD grand coalition): 2005-2009
  "Germany",       "CDU/CSU",   as.Date("2005-11-22"), as.Date("2009-10-27"),
  "Germany",       "SPD",       as.Date("2005-11-22"), as.Date("2009-10-27"),
  # Merkel II (CDU/CSU + FDP): 2009-2013
  "Germany",       "CDU/CSU",   as.Date("2009-10-28"), as.Date("2013-12-16"),
  "Germany",       "FDP",       as.Date("2009-10-28"), as.Date("2013-12-16"),
  # Merkel III (CDU/CSU + SPD grand coalition): 2013-2015
  "Germany",       "CDU/CSU",   as.Date("2013-12-17"), as.Date("2015-12-31"),
  "Germany",       "SPD",       as.Date("2013-12-17"), as.Date("2015-12-31"),

  # --- NETHERLANDS ---
  # Lubbers III (CDA + PvdA): 1989-1994
  "Netherlands",   "CDA",       as.Date("1990-01-01"), as.Date("1994-08-21"),
  "Netherlands",   "PvdA",      as.Date("1990-01-01"), as.Date("1994-08-21"),
  # Kok I (PvdA + VVD + D'66): 1994-1998
  "Netherlands",   "PvdA",      as.Date("1994-08-22"), as.Date("1998-08-02"),
  "Netherlands",   "VVD",       as.Date("1994-08-22"), as.Date("1998-08-02"),
  "Netherlands",   "D'66",      as.Date("1994-08-22"), as.Date("1998-08-02"),
  # Kok II (PvdA + VVD + D'66): 1998-2002
  "Netherlands",   "PvdA",      as.Date("1998-08-03"), as.Date("2002-07-21"),
  "Netherlands",   "VVD",       as.Date("1998-08-03"), as.Date("2002-07-21"),
  "Netherlands",   "D'66",      as.Date("1998-08-03"), as.Date("2002-07-21"),
  # Balkenende I (CDA + VVD + LPF): 2002-2003
  "Netherlands",   "CDA",       as.Date("2002-07-22"), as.Date("2003-05-26"),
  "Netherlands",   "VVD",       as.Date("2002-07-22"), as.Date("2003-05-26"),
  "Netherlands",   "LPF",       as.Date("2002-07-22"), as.Date("2003-05-26"),
  # Balkenende II (CDA + VVD + D'66): 2003-2006
  "Netherlands",   "CDA",       as.Date("2003-05-27"), as.Date("2006-07-06"),
  "Netherlands",   "VVD",       as.Date("2003-05-27"), as.Date("2006-07-06"),
  "Netherlands",   "D'66",      as.Date("2003-05-27"), as.Date("2006-07-06"),
  # Balkenende III (CDA + VVD caretaker): 2006-2007
  "Netherlands",   "CDA",       as.Date("2006-07-07"), as.Date("2007-02-21"),
  "Netherlands",   "VVD",       as.Date("2006-07-07"), as.Date("2007-02-21"),
  # Balkenende IV (CDA + PvdA + CU): 2007-2010
  "Netherlands",   "CDA",       as.Date("2007-02-22"), as.Date("2010-10-13"),
  "Netherlands",   "PvdA",      as.Date("2007-02-22"), as.Date("2010-10-13"),
  "Netherlands",   "CU",        as.Date("2007-02-22"), as.Date("2010-10-13"),
  # Rutte I (VVD + CDA minority): 2010-2012
  "Netherlands",   "VVD",       as.Date("2010-10-14"), as.Date("2012-11-04"),
  "Netherlands",   "CDA",       as.Date("2010-10-14"), as.Date("2012-11-04"),
  # Rutte II (VVD + PvdA): 2012-2015
  "Netherlands",   "VVD",       as.Date("2012-11-05"), as.Date("2015-12-31"),
  "Netherlands",   "PvdA",      as.Date("2012-11-05"), as.Date("2015-12-31"),

  # --- SPAIN ---
  "Spain",         "PSOE",      as.Date("1990-01-01"), as.Date("1996-05-05"),
  "Spain",         "PP",        as.Date("1996-05-06"), as.Date("2004-04-16"),
  "Spain",         "PSOE",      as.Date("2004-04-17"), as.Date("2011-12-20"),
  "Spain",         "PP",        as.Date("2011-12-21"), as.Date("2015-12-31"),

  # --- SWEDEN ---
  "Sweden",        "SAP",       as.Date("1990-01-01"), as.Date("1991-10-03"),
  # Bildt coalition (MSP + CP + Kd + FP): 1991-1994
  "Sweden",        "MSP",       as.Date("1991-10-04"), as.Date("1994-10-06"),
  "Sweden",        "CP",        as.Date("1991-10-04"), as.Date("1994-10-06"),
  "Sweden",        "Kd",        as.Date("1991-10-04"), as.Date("1994-10-06"),
  "Sweden",        "FP",        as.Date("1991-10-04"), as.Date("1994-10-06"),
  # Carlsson III + Persson (SAP minority): 1994-2006
  "Sweden",        "SAP",       as.Date("1994-10-07"), as.Date("2006-10-05"),
  # Reinfeldt (MSP + CP + Kd + FP): 2006-2014
  "Sweden",        "MSP",       as.Date("2006-10-06"), as.Date("2014-10-02"),
  "Sweden",        "CP",        as.Date("2006-10-06"), as.Date("2014-10-02"),
  "Sweden",        "Kd",        as.Date("2006-10-06"), as.Date("2014-10-02"),
  "Sweden",        "FP",        as.Date("2006-10-06"), as.Date("2014-10-02"),
  # Loefven (SAP minority): 2014-
  "Sweden",        "SAP",       as.Date("2014-10-03"), as.Date("2015-12-31"),

  # --- UNITED KINGDOM ---
  "United Kingdom","Conservatives", as.Date("1990-01-01"), as.Date("1997-05-01"),
  "United Kingdom","Labour",    as.Date("1997-05-02"), as.Date("2010-05-10"),
  # Cameron coalition (Con + LibDem): 2010-2015
  "United Kingdom","Conservatives", as.Date("2010-05-11"), as.Date("2015-12-31"),
  "United Kingdom","LibDems",   as.Date("2010-05-11"), as.Date("2015-12-31"),

  # --- DENMARK (for Party Congresses) ---
  "Denmark",       "SD",        as.Date("1993-01-25"), as.Date("2001-11-26"),
  "Denmark",       "V",         as.Date("2001-11-27"), as.Date("2011-10-02"),
  "Denmark",       "SD",        as.Date("2011-10-03"), as.Date("2015-06-27"),
  "Denmark",       "V",         as.Date("2015-06-28"), as.Date("2020-12-31")
)

# --- Apply governing party dummy to ParlSpeech ---
parlspeech_ideology <- parlspeech_ideology |>
  mutate(speech_date = as.Date(date)) |>
  left_join(gov_periods |> rename(gov_party = party), by = "countryname") |>
  mutate(in_gov = as.integer(
    speech_date >= gov_start & speech_date <= gov_end & party == gov_party
  )) |>
  group_by(speaker, date) |>
  slice_max(in_gov, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(-gov_start, -gov_end, -gov_party, -speech_date)

# --- Apply governing party dummy to EUSpeech ---
euspeech_ideology <- euspeech_ideology |>
  mutate(speech_date = as.Date(date, format = "%d-%m-%Y")) |>
  left_join(gov_periods |> rename(gov_party = party),
            by = c("country" = "countryname")) |>
  mutate(in_gov = case_when(
    is.na(gov_party) ~ 0L,
    speech_date >= gov_start & speech_date <= gov_end &
      partyabbrev == gov_party ~ 1L,
    TRUE ~ 0L
  )) |>
  group_by(speaker, date) |>
  slice_max(in_gov, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(-gov_start, -gov_end, -gov_party, -speech_date)

# --- Apply governing party dummy to Party Congresses ---
partycong_ideology <- partycong_ideology |>
  mutate(speech_year_int = as.integer(year)) |>
  left_join(gov_periods |> rename(gov_party = party), by = "countryname") |>
  mutate(
    gov_start_year = year(gov_start),
    gov_end_year   = year(gov_end),
    in_gov = as.integer(
      speech_year_int >= gov_start_year &
      speech_year_int <= gov_end_year &
      party == gov_party
    )
  ) |>
  group_by(speaker, year) |>
  slice_max(in_gov, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(-gov_start, -gov_end, -gov_party, -gov_start_year,
         -gov_end_year, -speech_year_int)

```

---

## Final Assembly

We standardize column names across all three datasets, normalize Flesch-Kincaid
scores within each dataset (as done in the paper), bind them together, and split
into the 8 country-level subsets used in the analysis. Year is zero-based
(relative to the earliest year in each subset).

```{r final-assembly}

# --- Standardize columns ---
parlspeech_final <- parlspeech_ideology |>
  mutate(dataset = "parlspeech") |>
  rename(country = countryname) |>
  select(speech_year, speech_month, text, speaker, party, partyname,
         parliament, country, lib_con_scale, left_right_scale,
         flesch_kincaid, in_gov, dataset)

euspeech_final <- euspeech_ideology |>
  mutate(dataset = "euspeech") |>
  rename(text = texts, partyname1 = party, partyabbrev1 = partyabbrev) |>
  select(speech_year, speech_month, text, speaker, country,
         lib_con_scale, left_right_scale, flesch_kincaid, in_gov,
         partyname1, partyabbrev1, dataset) |>
  rename(partyname = partyname1, party = partyabbrev1) |>
  select(speech_year, speech_month, text, speaker, party, partyname,
         country, lib_con_scale, left_right_scale, flesch_kincaid,
         in_gov, dataset)

partycong_final <- partycong_ideology |>
  mutate(dataset = "partycong") |>
  rename(text = texts, partyname1 = party_name, country1 = countryname) |>
  select(speech_year, text, speaker, party, partyname1, country1,
         lib_con_scale, left_right_scale, flesch_kincaid, in_gov, dataset) |>
  rename(partyname = partyname1, country = country1) |>
  select(speech_year, text, speaker, party, partyname, country,
         lib_con_scale, left_right_scale, flesch_kincaid, in_gov, dataset)

# --- Keep unstandardized FK for Figure 1, then standardize within dataset ---
parlspeech_final <- parlspeech_final |>
  mutate(flesch_kincaid_unstand = flesch_kincaid)
euspeech_final   <- euspeech_final |>
  mutate(flesch_kincaid_unstand = flesch_kincaid)
partycong_final  <- partycong_final |>
  mutate(flesch_kincaid_unstand = flesch_kincaid)

# Standardize (z-score) FK within each dataset, as done in the paper
parlspeech_final <- parlspeech_final |>
  mutate(flesch_kincaid = scale(flesch_kincaid)[, 1])
euspeech_final   <- euspeech_final |>
  mutate(flesch_kincaid = scale(flesch_kincaid)[, 1])
partycong_final  <- partycong_final |>
  mutate(flesch_kincaid = scale(flesch_kincaid)[, 1])

# --- Merge into one final dataframe ---
final_data <- bind_rows(parlspeech_final, euspeech_final, partycong_final)

# --- Split into 8 country-level subsets for analysis ---
dnk      <- final_data |> filter(dataset == "partycong",  country == "Denmark")
nld_cong <- final_data |> filter(dataset == "partycong",  country == "Netherlands")
ger      <- final_data |> filter(dataset == "parlspeech", country == "Germany")
esp      <- final_data |> filter(dataset == "parlspeech", country == "Spain")
gbr      <- final_data |> filter(dataset == "parlspeech", country == "United Kingdom")
swe      <- final_data |> filter(dataset == "parlspeech", country == "Sweden")
nld_parl <- final_data |> filter(dataset == "parlspeech", country == "Netherlands")
eus      <- final_data |> filter(dataset == "euspeech")

# Zero-base the year variable within each subset
dnk      <- dnk      |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
nld_cong <- nld_cong |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
ger      <- ger      |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
esp      <- esp      |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
gbr      <- gbr      |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
swe      <- swe      |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
nld_parl <- nld_parl |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))
eus      <- eus      |> mutate(speech_year = speech_year - min(speech_year, na.rm = TRUE))

# --- Save final cleaned data, makes sure it's full output---
if (n_sam == Inf) {
  save(final_data, dnk, nld_cong, ger, esp, gbr, swe, nld_parl, eus,
       file = "data/cleaned/cole_final_cleaned_data.RData")
  message("Final dataset saved. Total observations: ", nrow(final_data))
}



```
